---
title: "Betas"
author: "Jiaxi Li"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Load Clean Data

Load All.csv and select the applicable variables: log premiums, PERMNO, and isREIT

```{r}
# load useful packages
library(tidyverse)
library(here)
library(dplyr)
```

```{r, cache = TRUE}
# load data to include only log premiums, PERMNO, and isREIT
Reg <- read_csv(here("data/All.csv"))  %>%
  subset(select = c(1, 12:18))
```

# Beta Calculations with at least 3 years data

First, I would only choose the ones with more than 3 years worth of data (at least 36 data points) and also calculate the average risk premium for them.

```{r}
# 3 years (36 months)
k = 36
# Remove the empty rp's
Reg1 <- Reg %>%
  subset(!is.na(rp))

# there seems to be no empty rp so proceed with Reg
Returns <- Reg %>%
  group_by(PERMNO, isREIT) %>%
  summarise(n = n(),
            avg_rp = mean(rp))

# change isREIT to factor names
Reg$isREIT <- factor(Reg$isREIT, labels = c("non-REITs","REITs"))
Returns$isREIT <- factor(Returns$isREIT, labels = c("non-REITs","REITs"))

Simple_Beta1 <- function(Returns, Reg, k) {
  # This function simply extract the regression data with >= k data points
  # The primary inputs are:
  # 1. Returns: tibbles with "PERMNO", "isREIT", "n", "avg_rp".
  # 2. Reg: tibbles with "PERMNO", "rp", factors, and "isREIT"
  # 3. k: number of months we at least need for estimation
  
  # select the ones with more than l = k-1 data points
  l = k-1
  
  # select the data
  Reg2 <- Reg %>%
    full_join(Returns %>%
              subset(select = - 4)) %>%
    subset(n>l) %>%
    subset(select = - (ncol(Reg)+1))
    
  return(Reg2)
}

# Apply the function to extract data
Reg2 <- Simple_Beta1(Returns, Reg, k)

# summarize the data points
Reg2 %>%
  group_by(isREIT) %>%
  summarise(days_stocks = n())
  
Reg2 %>%
  group_by(isREIT, PERMNO) %>%
  summarise() %>%
  ungroup() %>%
  group_by(isREIT) %>%
  summarise(stocks = n())
```

There are 1,496,878 usable non-REITs day&returns and 49,055 usable REITs day&returns. They are from 11,503 non-REITs stocks and 361 REITs stocks.

I will apply all the available data for each asset to calculate the asset beta here. The model used is Fama French 5 Factor model.

```{r, eval= FALSE}
# This part should theoretically run, but it was taking too much time (maybe a better computer can do it).
# group PERMNO and isREIT and then do the regression (Reg2 is the result of previous function)
Reg1 <- Reg2 %>%
  group_by(PERMNO, isREIT)

temp <- Reg1 %>%
  summarise(betas = list(lm(rp ~ .- PERMNO - isREIT, data = Reg1) %>%
                           coefficients()))

# extract the betas
Betas <- do.call(rbind.data.frame, temp$betas) %>%
              as_tibble()

# rename the beta variables
names(Betas) = names(temp$betas[[1]])

# join the identifiers and betas
Betas1 <- cbind(temp[, c(1,2)], Betas) %>%
  as_tibble()
```

According to https://nelsonareal.net/blog/2017/06/speeding_up_ols.html, I will use speedlm to do the regression since it is the fastest with formula. It is from the "speedglm" package. Since group_by took too long to run, I have a for loop instead.

```{r, cache = TRUE}
Simple_Beta2 <- function(Returns,Reg2,k){
  # This function evaluate the betas with the factors
  # The primary inputs are:
  # 1. Returns: tibbles with "PERMNO", "isREIT", "n", "avg_rp".
  # 2. Reg2: tibbles result from Simple_Beta1, with "PERMNO", "rp", factors, and "isREIT"
  # 3. k: number of months we at least need for estimation
  
  # select the ones with more than l = k-1 data points
  l = k-1

  # extract all the stocks with more than l data points
  Returns1<- Returns %>%
  subset(n>l)

  # group PERMNO and isREIT and then do the regression
  # presave the result
  Betas <- Reg2 %>%
    subset(select = -2) %>%
    group_by(PERMNO, isREIT) %>%
    summarise_all(funs(mean)) %>%
    rename_all(funs(toupper)) %>%
    rename(isREIT = ISREIT)
  # use the for-loop to calculate the betas without intercept
  for (i in 1:nrow(Returns1)) {
    Betas[i, 3:7] <- (speedglm::speedlm(rp ~ .- PERMNO - isREIT -1, data = Reg1 %>%
                                    subset(PERMNO == Returns1$PERMNO[i])) %>%
                        coefficients())
  }
  return(Betas)
}

# Find the Betas with function
Betas_36 <- Simple_Beta2(Returns,Reg2,k)
```

Save all the beta result in the csv file.
```{r, eval = FALSE}
write_csv(Betas_36, here("data","Betas_36.csv"))
```

Since the second step should be a multi-linear regression, graph each beta against return may not result in a straight line. Instead, the distribution is presented here:
```{r}
# assign Betas
Betas <- Betas_36

# fix Betas1 for the plot
Betas <- Betas %>%
  gather(types, betas, -c("PERMNO", "isREIT"))

# fix the order of factors
Betas$types <- factor(Betas$types, levels = c("MKT_RF","SMB","HML","RMW","CMA"))

# plot the betas
Betas %>%
  group_by(isREIT, types, betas) %>%
  ggplot(aes(types, betas)) +
  geom_violin() +
  geom_boxplot() +
  labs(title= paste("Estimated Betas for the Fama French 5 Factor Model with k=", toString(k),""),
       caption = "Source: CRSP, Ken French Data Library, and Ziman",
       x = "Factors", y = "Beta") + 
  facet_wrap(~ isREIT)+
  theme(plot.title = element_text(hjust=0.5))
```

With the distribution plot presented above, I also include the quartiles in the table below.
```{r}
# find the quantiles as well as the mean for each beta
Beta_stats <- Betas %>%
  group_by(isREIT, types) %>%
  summarise(mean = mean(betas),
            variance = var(betas),
            kurtosis = moments::kurtosis(betas),
            min = min(betas),
            Q1 = quantile(betas, probs = 0.25),
            median = median(betas),
            Q3 = quantile(betas, probs = 0.75),
            max = max(betas)) %>%
  gather(stat,  values, -c("isREIT","types")) %>%
  spread(types, values)

# fix the order of the table
Beta_stats <- Beta_stats %>%
  arrange(match(stat, c("mean","variance","kurtosis","min","Q1","median","Q3","max"))) %>%
  arrange_at(1)

# present the table
Beta_stats %>%
  knitr::kable(format = 'html',
        caption = paste("Statistics of Betas of the Fama French 5 Factor Model with k=", toString(k),""))
```

It seems that the distribution for REITs betas and non-REITs betas are similar.

# Beta Calculations with at least 4 years data

Second, I would only choose the ones with more than 4 years worth of data (at least 48 data points) and also analysis the same way as before.

```{r}
# 4 years is 48 months
k = 48

# Apply the function to extract data
Reg2 <- Simple_Beta1(Returns, Reg, k)

# summarize the data points
Reg2 %>%
  group_by(isREIT) %>%
  summarise(days_stocks = n())
  
Reg2 %>%
  group_by(isREIT, PERMNO) %>%
  summarise() %>%
  ungroup() %>%
  group_by(isREIT) %>%
  summarise(stocks = n())
```

There are 1,438,238 usable non-REITs day&returns and 47,098 usable REITs day&returns. They are from 10,085 non-REITs stocks and 314 REITs stocks.

I will apply all the available data for each asset to calculate the asset beta here. The model used is Fama French 5 Factor model.

```{r, cache = TRUE}
# Find the Betas with function
Betas_48 <- Simple_Beta2(Returns,Reg2,k)
```

Save all the beta result in the csv file.
```{r, eval = FALSE}
write_csv(Betas_48, here("data","Betas_48.csv"))
```

The distribution, again, is presented here:
```{r}
# assign Betas
Betas <- Betas_48

# fix Betas1 for the plot
Betas <- Betas %>%
  gather(types, betas, -c("PERMNO", "isREIT"))

# fix the order of factors
Betas$types <- factor(Betas$types, levels = c("MKT_RF","SMB","HML","RMW","CMA"))

# plot the betas
Betas %>%
  group_by(isREIT, types, betas) %>%
  ggplot(aes(types, betas)) +
  geom_violin() +
  geom_boxplot() +
  labs(title= paste("Estimated Betas for the Fama French 5 Factor Model with k=", toString(k),""),
       caption = "Source: CRSP, Ken French Data Library, and Ziman",
       x = "Factors", y = "Beta") + 
  facet_wrap(~ isREIT)+
  theme(plot.title = element_text(hjust=0.5))
```

With the distribution plot presented above, I also include the quartiles in the table below.
```{r}
# find the quantiles as well as the mean for each beta
Beta_stats <- Betas %>%
  group_by(isREIT, types) %>%
  summarise(mean = mean(betas),
            variance = var(betas),
            kurtosis = moments::kurtosis(betas),
            min = min(betas),
            Q1 = quantile(betas, probs = 0.25),
            median = median(betas),
            Q3 = quantile(betas, probs = 0.75),
            max = max(betas)) %>%
  gather(stat,  values, -c("isREIT","types")) %>%
  spread(types, values)

# fix the order of the table
Beta_stats <- Beta_stats %>%
  arrange(match(stat, c("mean","variance","kurtosis","min","Q1","median","Q3","max"))) %>%
  arrange_at(1)

# present the table
Beta_stats %>%
  knitr::kable(format = 'html',
        caption = paste("Statistics of Betas of the Fama French 5 Factor Model with k=", toString(k),""))
```

It seems that the distribution for REITs betas and non-REITs betas are similar.

# Beta Calculations with at least 5 years data

Second, I would only choose the ones with more than 5 years worth of data (at least 60 data points) and also analysis the same way as before.

```{r}
# 5 years is 60 months
k = 60

# Apply the function to extract data
Reg2 <- Simple_Beta1(Returns, Reg, k)

# summarize the data points
Reg2 %>%
  group_by(isREIT) %>%
  summarise(days_stocks = n())
  
Reg2 %>%
  group_by(isREIT, PERMNO) %>%
  summarise() %>%
  ungroup() %>%
  group_by(isREIT) %>%
  summarise(stocks = n())
```

There are 1,364,749 usable non-REITs day&returns and 44,560 usable REITs day&returns. They are from 8,714 non-REITs stocks and 266 REITs stocks.

I will apply all the available data for each asset to calculate the asset beta here. The model used is Fama French 5 Factor model.

```{r, cache = TRUE}
# Find the Betas with function
Betas_60 <- Simple_Beta2(Returns,Reg2,k)
```

Save all the beta result in the csv file.
```{r, eval = FALSE}
write_csv(Betas_60, here("data","Betas_60.csv"))
```

The distribution, again, is presented here:
```{r}
# assign Betas
Betas <- Betas_60

# fix Betas1 for the plot
Betas <- Betas %>%
  gather(types, betas, -c("PERMNO", "isREIT"))

# fix the order of factors
Betas$types <- factor(Betas$types, levels = c("MKT_RF","SMB","HML","RMW","CMA"))

# plot the betas
Betas %>%
  group_by(isREIT, types, betas) %>%
  ggplot(aes(types, betas)) +
  geom_violin() +
  geom_boxplot() +
  labs(title= paste("Estimated Betas for the Fama French 5 Factor Model with k=", toString(k),""),
       caption = "Source: CRSP, Ken French Data Library, and Ziman",
       x = "Factors", y = "Beta") + 
  facet_wrap(~ isREIT)+
  theme(plot.title = element_text(hjust=0.5))
```

With the distribution plot presented above, I also include the quartiles in the table below.
```{r}
# find the quantiles as well as the mean for each beta
Beta_stats <- Betas %>%
  group_by(isREIT, types) %>%
  summarise(mean = mean(betas),
            variance = var(betas),
            kurtosis = moments::kurtosis(betas),
            min = min(betas),
            Q1 = quantile(betas, probs = 0.25),
            median = median(betas),
            Q3 = quantile(betas, probs = 0.75),
            max = max(betas)) %>%
  gather(stat,  values, -c("isREIT","types")) %>%
  spread(types, values)

# fix the order of the table
Beta_stats <- Beta_stats %>%
  arrange(match(stat, c("mean","variance","kurtosis","min","Q1","median","Q3","max"))) %>%
  arrange_at(1)

# present the table
Beta_stats %>%
  knitr::kable(format = 'html',
        caption = paste("Statistics of Betas of the Fama French 5 Factor Model with k=", toString(k),""))
```

It seems that the distribution for REITs betas and non-REITs betas are similar.

## Number of REITs at each month
```{r, cache = TRUE}
# Number of REITs each time
All <- read_csv(here("data/All.csv"))

All %>%
  subset(select = c(1, 2,18)) %>%
  subset(isREIT == TRUE) %>%
  group_by(date) %>%
  summarise(n = n()) %>%
  ggplot(aes(date, n)) +
  geom_line() +
  labs(title= "Evolution of Number of REITs",
       caption = "Source: Ken French Data Library and Ziman",
       x = "Date", y = "Number of REITs") + 
  theme(plot.title = element_text(hjust=0.5))
```

One way to tackle the unbalanced panel problem is to pick the stocks to form a balanced panel. However, this would result in a survivorship bias. I will check the IV-GMM method for alternative solution. 