---
title: "Betas"
author: "Jiaxi Li"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Load Clean Data

Load All.csv and select the applicable variables: log premiums, PERMNO, and isREIT

```{r}
# load useful packages
library(tidyverse)
library(here)
library(dplyr)

# load data to include only log premiums, PERMNO, and isREIT
Reg <- read_csv(here("data/All.csv"))  %>%
  subset(select = c(1, 12:18))
```

# Beta Calculations

First, I would only choose the ones with more than 3 years worth of data (at least 36 data points) and also calculate the average risk premium for them.

```{r, cache = TRUE}
# Remove the empty rp's
Reg1 <- Reg %>%
  subset(!is.na(rp))

# there seems to be no empty rp so proceed with Reg
Returns <- Reg %>%
  group_by(PERMNO, isREIT) %>%
  summarise(n = n(),
            avg_rp = mean(rp))

# select the ones with more than 35 data points
Reg2 <- Reg %>%
  full_join(Returns %>%
              subset(select = - 4)) %>%
  subset(n>35) %>%
  subset(select = - 9)

# summarize the data points
Reg2 %>%
  group_by(isREIT) %>%
  summarise(n = n())

Reg2 %>%
  group_by(isREIT, PERMNO) %>%
  summarise() %>%
  ungroup() %>%
  group_by(isREIT) %>%
  summarise(n = n())
```

There are 1,496,878 usable non-REITs day&returns and 49,055 usable REITs day&returns. They are from 11,503 non-REITs stocks and 361 REITs stocks.

I will apply all the available data for each asset to calculate the asset beta here. The model used is Fama French 5 Factor model.

```{r, eval= FALSE}
# This part should theoretically run, but it was taking too much time (maybe a better computer can do it).
# group PERMNO and isREIT and then do the regression
Reg1 <- Reg2 %>%
  group_by(PERMNO, isREIT)

temp <- Reg1 %>%
  summarise(betas = list(lm(rp ~ .- PERMNO - isREIT, data = Reg1) %>%
                           coefficients()))

# extract the betas
Betas <- do.call(rbind.data.frame, temp$betas) %>%
              as_tibble()

# rename the beta variables
names(Betas) = names(temp$betas[[1]])

# join the identifiers and betas
Betas1 <- cbind(temp[, c(1,2)], Betas) %>%
  as_tibble()
```

According to https://nelsonareal.net/blog/2017/06/speeding_up_ols.html, I will use speedlm to do the regression since it is the fastest with formula. It is from the "speedglm" package. Since group_by took too long to run, I have a for loop instead.

```{r, cache = TRUE}
# load speedglm for fast regression
library(speedglm)

# extract all the stocks with more than 36 data points
Returns_36 <- Returns %>%
  subset(n>35)

# group PERMNO and isREIT and then do the regression
# presave the result
Betas <- Reg2 %>%
  group_by(PERMNO, isREIT) %>%
  summarise() %>%
  mutate(b_m = 0,
         b_smb = 0,
         b_hml = 0,
         b_rmw = 0,
         b_cma = 0)

# use the for-loop to calculate the betas without intercept
for (i in 1:nrow(Returns_36)) {
  Betas[i, 3:7] <- (speedlm(rp ~ .- PERMNO - isREIT -1, data = Reg1 %>%
                                  subset(PERMNO == Returns_36$PERMNO[i])) %>%
                      coefficients())
}

```

Save all the beta result in the csv file.
```{r, eval = FALSE}
 write_csv(Betas, here("data","Betas.csv"))
```

Since the second step should be a multi-linear regression, graph each beta against return may not result in a straight line. Instead, the distribution is presented here.
```{r}
# change isREIT to factor names
Betas1 <- Betas
Betas1$isREIT <- factor(Betas1$isREIT, labels = c("non-REITs","REITs"))

# fix Betas1 for the plot 
Betas1 <- Betas1 %>%
  gather(types, betas, -c("PERMNO", "isREIT")) %>%
  mutate(types = toupper(types %>%
                           str_replace('b_', "")))

# fix the order of factors
Betas1$types <- factor(Betas1$types, levels = c("M","SMB","HML","RMW","CMA"))

# plot the betas
Betas1 %>%
  group_by(isREIT, types, betas) %>%
  ggplot(aes(types, betas)) +
  geom_violin() +
  geom_boxplot() +
  labs(title= "Estimated Betas for the Fama French 5 Factor Model",
       caption = "Source: CRSP, Ken French Data Library, and Ziman",
       x = "Factors", y = "Beta") + 
  facet_wrap(~ isREIT)+
  theme(plot.title = element_text(hjust=0.5))
```

With the distribution plot presented above, I also include the quartiles in the table below.
```{r}
# find the quantiles as well as the mean for each beta
Beta_stats <- Betas1 %>%
  group_by(isREIT, types) %>%
  summarise(mean = mean(betas),
            variance = var(betas),
            kurtosis = moments::kurtosis(betas),
            min = min(betas),
            Q1 = quantile(betas, probs = 0.25),
            median = median(betas),
            Q3 = quantile(betas, probs = 0.75),
            max = max(betas)) %>%
  gather(stat,  values, -c("isREIT","types")) %>%
  spread(types, values)

# fix the order of the table
Beta_stats <- Beta_stats %>%
  arrange(match(stat, c("mean","variance","kurtosis","min","Q1","median","Q3","max"))) %>%
  arrange_at(1)

# present the table
Beta_stats %>%
    knitr::kable(format = 'html',
        caption = 'Statistics of Betas of the Fama French 5 Factor Model')

```

It seems that the distribution for REITs betas and non-REITs betas are similar.