---
title: "Betas"
author: "Jiaxi Li"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Load Clean Data

Load All.csv and select the applicable variables: log premiums, PERMNO, and isREIT

```{r}
# load useful packages
library(tidyverse)
library(here)
library(dplyr)

# load data to include only log premiums, PERMNO, and isREIT
Reg <- read_csv(here("data/All.csv"))  %>%
  subset(select = c(1, 12:18))
```

# Beta Calculations

First, I would only choose the ones with more than 3 years worth of data (at least 36 data points) and also calculate the average risk premium for them.

```{r}
# Remove the empty rp's
Reg1 <- Reg %>%
  subset(!is.na(rp))

# there seems to be no empty rp so proceed with Reg
Returns <- Reg %>%
  group_by(PERMNO, isREIT) %>%
  summarise(n = n(),
            avg_rp = mean(rp))

# select the ones with more than 35 data points
Reg2 <- Reg %>%
  full_join(Returns %>%
              subset(select = - 4)) %>%
  subset(n>35) %>%
  subset(select = - 9)

# summarize the data points
Reg2 %>%
  group_by(isREIT) %>%
  summarise(n = n())

Reg2 %>%
  group_by(isREIT, PERMNO) %>%
  summarise() %>%
  ungroup() %>%
  group_by(isREIT) %>%
  summarise(n = n())
```

There are 1,496,878 usable non-REITs day&returns and 49,055 usable REITs day&returns. They are from 11,503 non-REITs stocks and 361 REITs stocks.

I will apply all the available data for each asset to calculate the asset beta here. The model used is Fama French 5 Factor model.

```{r}
# group PERMNO and isREIT and then do the regression
Reg1 <- Reg %>%
  group_by(PERMNO, isREIT)

temp <- Reg1 %>%
  summarise(betas = list(lm(rp ~ .- PERMNO - isREIT, data = Reg1) %>%
                           coefficients()))

# extract the betas
Betas <- do.call(rbind.data.frame, temp$betas) %>%
              as_tibble()

# rename the beta variables
names(Betas) = names(temp$betas[[1]])

# join the identifiers and betas
Betas1 <- cbind(temp[, c(1,2)], Betas) %>%
  as_tibble()

```

