---
title: "Cleaning"
author: "Jiaxi Li"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

In this file, I first load all the data. After transforming simple returns into log returns, I remove the irrelevant variables and save all the data into a data file in the data folder.

# Load Data

```{r}
# load useful packages
library(tidyverse)
library(here)
library(dplyr)

# load the data and rename to make common names the same; only keep the REIT ones with share code 18 and exchange code above 0

Regular <- read_csv(here("data/Regular.csv"))

REIT <- readxl::read_xlsx(here("data/REIT.xlsx")) %>%
  # rename the variables as names in Regular
  rename(PERMNO = "CRSP Permanent Security Identifier",
         date = "Calendar Date",
         SHRCD = "Share Code",
         EXCHCD = "Exchange Code",
         RET = "Used Price Total Return...27",
         cap = "Market Capitalization",
         REITType = "REIT Type",
         PropertyType = "Property Type") %>%
  # keep share code = 18 and exchange code > 0
  filter(SHRCD == 18,
         EXCHCD > 0)

Factors <- read_csv(here("data/FF5.csv")) %>%
  # rename the date and Mkt_RF
  rename(date = "X1",
         Mkt_RF = "Mkt-RF") %>%
  # change all returns into decimals
  mutate_at(vars(-date), function(x, na.rm = FALSE) (x/100))
```


# Clean Regular Stocks

The selected "Regular" Stocks are the ones with share code 10 and 11.

```{r}
# Generate lagged capitalization for potential rebalancing and change date into yyyymm. Keep exchange code and share code. Note: RET is simple return.
Regular1 <- Regular %>%
  group_by(PERMNO) %>%
  mutate(date = floor(date/100),
         cap = abs(PRC) * SHROUT,
         lagcap = lag(cap, n = 1)) %>%
  ungroup() %>%
  subset(select = -c(5,7,8))
```

# Clean REIT Stocks

The selected "REIT" Stocks are the ones with share code 18.

```{r}
# Generate lagged capitalization for potential rebalancing and change date into yyyymm. Keep exchange code and share code. Note: RET is simple return.
REIT1 <- REIT %>%
  group_by(PERMNO) %>%
  # change date format
  mutate(date = gsub("[^[:digit:]]", "", date),
         date = floor(as.numeric(date)/100),
         lagcap = lag(cap, n = 1)) %>%
  ungroup() %>%
  subset(select = names(Regular1))
```

# Join all three dataset together

Join all data, remove non-existing returns

```{r}
# bind Regular1 and REIT1, then join Factors
All <- bind_rows(Regular1, REIT1) %>%
  inner_join(Factors) %>%
  # create risk premium RP (simple) and rp (log), and all log factors premium
  mutate(RP = RET - RF,
         rp = log(1 + RP),
         mkt_rf = log(1 + Mkt_RF),
         smb = log(1 + SMB),
         hml = log(1 + HML),
         rmw = log(1 + RMW),
         cma = log(1 + CMA),
         isREIT = (SHRCD == 18)) %>%
  # remove share code and exchange code
  subset(select = -c(3,4)) %>%
  filter(!is.na(rp))
```

# Verify the non-REITs

There are some equities are labeled both REITs and non-REITs. After investigating a few of them, it seems that CRSP first label the REITs as common stocks when it is just listed and then change it as REITs later. I would correct them here.

```{r}
# group PERMNO and isREIT and check those PERMNOs classified as both
All_1 <- All %>%
  group_by(PERMNO, isREIT) %>%
  summarise()

# check whether the PERMNOs are unique here (there are ones that is labeled both REITs and non-REITs)
Problems <- All_1 %>%
  ungroup() %>%
  group_by(PERMNO) %>%
  summarise(n = n()) %>%
  filter(n > 1) %>%
  inner_join(All_1)

# It seems one with both labels are all REITs (I will trust the REITs data), and change all them to REITs
All <- All %>%
  mutate(isREIT = (isREIT|(PERMNO %in% Problems$PERMNO)))

# Remove duplicate ones
All <- All %>%
  unique()

# check whether there are only one PERMNOs (return) each day. For those with a problem, I would trust the CRSP data. I checked the CRSP data, and it seems the first ones are accurate.

All_2 <- All %>%
  group_by(PERMNO, date) %>%
  summarise()

Day_problem <- All %>%
  group_by(PERMNO, date) %>%
  summarise(n = n()) %>%
  filter(n > 1) %>%
  inner_join(All)

# Remove the repeating ones that are inaccurate (even ones in Day_problem)
Day_problem1 <- Day_problem %>%
  subset(1:nrow(Day_problem) %in% seq(2,nrow(Day_problem), by = 2))

All_3 <- Day_problem1 %>%
  full_join(All)
All_3 <- All_3 %>%
  subset(is.na(n)) %>%
  subset(select = -3)

All_3 %>%
  group_by(isREIT) %>%
  summarise(n = n())
```

Then, I will extract a table with desired REITs Property types vs. PERMNO. We would want only equity REITs (REIT TYPE == 2) and leave out anything with a property type = 0, 1, 6, 7, or 8.  That gets us away from REITs where we don’t know what they are actually doing (0 or 1), are mortgage REITs (6 or 7), or in residential (8) which is brand new or incredibly unusual.

```{r}
# Select Equity REIT and PropertyType 
Properties <- REIT %>%
  filter(REITType == 2,
         PropertyType %in% c(2:5,9,10)) %>%
  group_by(PERMNO, PropertyType) %>%
  summarise(n = n())

Prop <- Properties %>%
  subset(select = - n) %>%
  ungroup() %>%
  group_by(PERMNO) %>%
  summarise(n = n()) %>%
  filter(n > 1)

# Remove the ones only appears once
Properties <- Properties %>%
  filter(!((n == 1) & (PERMNO %in% Prop$PERMNO))) %>%
  subset(select = - n)
```

Finally, we have matched 1,625,850 day&returns. There are 50,885 day&returns for REITs and 1,574,965 day&returns for non-REITs.

# Save all the data
```{r, eval = FALSE}
# Save file as a csv
 write_csv(All_3, here("data","All.csv"))
 write_csv(Problems, here("data","Problems.csv"))
 write_csv(Day_problem, here("data","Day_problem.csv"))
 write_csv(Properties, here("data","Properties.csv"))
```

